GHRVis:

### 2022-02-18 paper/section/data.tex
draft: update

Obtaining the data used in this paper is challenging, especially when an Electronic Health Records (EHR) dataset is involved \cite{wang2021EHRa}. The data acquisition involves multiple complicated steps. The first step is to obtain both geospaital data and EHR data that are compatible for projection. The second step is to pre-process the data for removing empty and erroneous values. The final step is to transform the data into a format that is suitable for visualization.

@@

Geospatial data, commonly known as shapefiles, were obtained from the following sources. After acquiring these data, we used QGIS \cite{qgisWelcome} to manually adjust projections and merge them into one shapefile. Finally, mapshaper \cite{blochMapshaper} is used to converted the merged shapefile into a TopoJSON \cite{TopoJSON} file to reduce the file size in order to improve the performance of our software.

@@

Clinical Commissioning Groups (CCGs) are the main geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}. The number of CCGs is changing each year due to the NHS annual re-organization, the most up-to-date shapefile is available from the Open Geography portalx \cite{opengeographyportalxOpen}. We decided to use the CCG shapefile from 2020 as at the time of writing, there is no EHR data published based on the latest CCG re-organization took place in April 2021.

@@

We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent and River Great Ouse in England. These rivers were chosen as they pass through regions with clusters of CCGs, and provide informative geographical and topological cues.

We first obtained a relation ID by searching a river, e.g. River Thames, on OpenStreetMap. The obtained relation ID was used to construct a query (See Listing~\ref{overpass}) which enabled use to download the entire river shapefile through Overpass Turbo \cite{overpassturboOverpass}.

\begin{lstlisting}[caption={The query that downloads the shapefile of River Thames from OpenStreetMap via Overpass Turbo API.}, label={overpass},captionpos=b]
    relation(2263653);>>;
    out skel;
\end{lstlisting}

>>>>

Obtaining the data is challenging, especially when an Electronic Health Records (EHR) dataset is involved \cite{wang2021EHRa}. The data acquisition involves multiple steps. The first step is to obtain both geospatial boundaries and EHR data that match one another. The second step is to pre-process the EHR data for removing empty and erroneous values. The final step is to transform the data into a format that is suitable for visualization.

@@

Geospatial boundaries, or shapefiles, were obtained from the following sources. After acquiring the shapefiles, we used QGIS \cite{qgisWelcome} to manually adjust projections and merge them into one shapefile. Finally, mapshaper \cite{blochMapshaper} is used to convert the merged shapefile into a TopoJSON \cite{TopoJSON} file to reduce the file size and improve the performance of our implementation.

{
\begin{figure}[tb!]
    \centering
    \includegraphics[width=0.6\columnwidth]{figure/ccg.png}
    \caption{A map of 106 CCGs in England as of April 2021, obtained from the Open Geography portalx \cite{opengeographyportalxOpen}.}
    \label{fig:ccg}
\end{figure}
}

@@

Clinical Commissioning Groups (CCGs) are the main geographic unit of the National Health Service (NHS) in the UK \cite{nhsNHS}. The number of CCGs changes each year due to the NHS re-organization. The most up-to-date shapefile is available from the Open Geography portalx \cite{opengeographyportalxOpen}. We decided to use the CCG shapefile from 2020 at the time of writing, there is no public EHR data published based on the latest CCG re-organization that took place in April 2021.

@@

We used OpenStreetMap \cite{openstreetmapRelation} as our data source to obtain shapefiles for River Thames, River Trent and River Great Ouse in England. These rivers were chosen as they pass through regions with dense population, and provide informative geographical and topological cues.

We first obtain a relation ID by searching for a river, e.g. River Thames, on OpenStreetMap. The relation ID is used to construct a query (See Listing~\ref{overpass}) which enables the user to download the entire river shapefile using Overpass Turbo \cite{overpassturboOverpass}.

\begin{lstlisting}[caption={The query that downloads the shapefile of River Thames from OpenStreetMap via the Overpass Turbo API.}, label={overpass},captionpos=b]
    relation(2263653);>>;
    out skel;
\end{lstlisting}

### 2022-02-18 paper/section/main.tex
draft: update

\section{\software}

\algoref{alg:UpdateNodePosition} describes the overview of \software. We first load all processed data from \secref{sec:{Data Description}} and apply VPSC \cite{dwyer2006fast} recursively to remove overlaps. We chose VPSC over other node overlap removal algorithms since VPSC is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}. During each recursion, we check node trajectories (See \algoref{alg:check river crossing}) and move back nodes that crossed a river. A recursion ends when 1) no overlaps are present; 2) no nodes crossed a river. We then increase the node size and repeat the recursion until the entire map canvas is filled.

@@

\begin{algorithm}[tb!]
    \caption{The procedure to update node positions by removing overlaps and prevent nodes from crossing rivers.}\label{alg:UpdateNodePosition}
    \textbf{Variables:} \\
    $NodeList \gets$ a list of nodes representing CCGs \\
    $Node_{size}\gets$ the current size of all nodes \\
    $Node_{inc} \gets$ the increment node size of each iteration \\
    $Node_{max} \gets$ the maximum size of a node \\
    $Node_{stale} \gets$ the maximum number of iterations before a stalemate \\
    \begin{algorithmic}[1]
        \Procedure{UpdateNodePosition}{}
        \While{$ Node_{size} < Node_{max} $}

        \State $ nodeList_{ora} \gets $ RunVPSC ($ NodeList $)

            \For{$ node^i_{ora} \in nodeList_{ora} $}

                \If {MoveNode ($ Node^i, node^i_{ora} $)}
                    \State $ cross \gets $ \Call{CheckRiverCrossing}{$ Node^i, node^i_{ora} $}

                    \If{$ cross = True $}
                        \State $ Node^i.stale \gets Node^i.stale + 1 $

                        \If{$ Node^i.stale < Node_{stale} $}
                            \State \Call{MoveNode}{$ Node^i $}
                        \Else
                            \State \Call{DeriveCorridor}{$ Node^i, node^i_{ora} $}
                        \EndIf

@@

        \State $Node_{size} \gets Node_{size} + Node_{inc}$

@@

\begin{algorithm}[tb!]
    \caption{The procedure to update a node's position.}\label{alg:move position}
    \begin{algorithmic}[1]
        \Procedure{MoveNode}{$ Node, Node_{new} $}
        \If{$ Node.x = Node_{new}.x $ and $ Node.y = Node_{new}.y $}
            \State \Return{$ False $}
        \EndIf

        \State init. $ x,y $

        \If{$ Node_{new}$ is $ undefined $}
            \State $previous \gets Node.history.pop() $
            \State $ x \gets previous.x,~y \gets previous.y $
        \Else
            \State $ x \gets Node_{new}.x,~y \gets Node_{new}.y $
            \State $ Node.history.push(x, y) $
        \EndIf

        \State $ Node.x \gets x $
        \State $ Node.y \gets y $
        \State \Return{$ True $}
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

@@

\subsection{River Cross Checking}

We use rivers as topological boundaries to prevent nodes from crossing them. When a node's position changes, we check if the node's trajectory intersects any segment of a river, see \algoref{alg:check river crossing}. A boundary collision detection is performed to reduce the number of line intersection detections required.

% \begin{noindent}
\begin{algorithm}[tb!]
    \caption{The procedure to check if a node crosses a river.}\label{alg:check river crossing}
    \textbf{Variables:} \\
    $RiverEdges \gets$ a list of river edges formed by two points \\

    \begin{algorithmic}[1]
        \Procedure{CheckRiverCrossing}{$ Node, Node_{ora} $}
            \For{$ edge \in RiverEdges $}
                \State $ node_b \gets $ GenerateBoundary ($ Node, Node_{ora} $)
                \State $ edge_b \gets $ GenerateBoundary ($ edge $)
                \State $ collide \gets $ DetectBoundaryCollision ($ node_b, edge_b $)
                \If{$ collide = True $}
                    \State $ node_l \gets $ GenerateLine ($ Node, Node_{ora} $)
                    \State $ edge_l \gets $ GenerateLine ($ edge $)
                    \State $ intersect \gets $ DetectLineIntersection ($ node_l, edge_l $)
                    \If{$ intersect = True $}
                        \State \Return{$ True $}
                    \EndIf
                \EndIf
            \EndFor
            \State \Return{$ False $}
        \EndProcedure
    \end{algorithmic}
\end{algorithm}
%\end{noindent}

@@

As the VPSC always tries to produce an optimal node layout where node spread and movement are minimized, a node might be repeatedly traveling between two positions due to the lack of available space, creating a stalemate situation for recursions, as shown in \figref{fig:stalemate}. In order to create space for VPSC to generate a new layout, we introduce a user-adjustable parameter, $ Node_{stale} $, to break the stalemate: if a node is traveling between two positions for more than $ Node_{stale} $ iterations, it is considered to be in a stalemate. A stalemate triggers a corridor derivation which moves all nodes within the corridor by $ \mathcal{D} $ set by the user (See \figref{fig:corridor} and \algoref{alg:derive corridor}).

{
\begin{figure}[tb!]
    \centering
    \includegraphics[width=\columnwidth]{figure/stalemate.png}
    \caption{A stalemate situation is when a node repeatedly travels between two positions (A and B) for more than $Node_{stale}$ iterations. }
    \label{fig:stalemate}
\end{figure}
}

{
\begin{figure}[tb!]
    \centering
    \includegraphics[width=\columnwidth]{figure/corridor.png}
    \caption{When a stalemate occurs, a corridor (black rectangular) is derived based on the current (A) and previous (B) positions of the node that crossed a river, as described in \algoref{alg:derive corridor}. All nodes within the corridor are moved by $ \mathcal{D} $ in the direction of $ \vec{BA} $. }
    \label{fig:corridor}
\end{figure}
}

@@

\begin{algorithm}[tb!]
    \caption{The procedure to derive a corridor to move included nodes. We use a SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending to the bottom, there is no negative axes.}\label{alg:derive corridor}
    \textbf{Variables:} \\
    $\mathcal{C}_l \gets$ the length of a corridor \\
    $\mathcal{C}_w \gets$ the width of a corridor \\
    $\mathcal{D} \gets$ the distance to move nodes within the corridor \\
    $Node_{size}\gets$ the current size of all nodes \\
    \begin{algorithmic}[1]
        \Procedure{DeriveCorridor}{$ Node, Node_{old} $}
        \State MoveNode ($ Node $)

        \State $ \Delta x \gets Node.x - Node_{old}.x,~ \Delta y \gets Node.y - Node_{old}.y $

@@

        \State $ pLine = $ \Call{DerivePoint}{$ \frac{-1}{ slope } $, $ Node $, $ \mathcal{C}_w $}
        \State $ sideLine_1 = $ \Call{DerivePoint}{$ slope $, $ pLine.a $, $ \mathcal{C}_l $, $ True $}
        \State $ sideLine_2 = $ \Call{DerivePoint}{$ slope $, $ pLine.b $, $ \mathcal{C}_l $, $ True $}
        \State $ corridor \gets
            \begin{bmatrix}
                pLine.a &
                pLine.b \\

                sideLine_1.a &
                sideLine_2.b \\
            \end{bmatrix} $

        \For{$ node^i $ inside $ corridor $}
            \State $ destination = $ \Call{DerivePoint}{$ \frac{-1}{ slope } $, $ node^i $, $ \mathcal{D} $}

            \State MoveNode ($ node^i, destination.b $)
        \EndFor

        \item[]

        \Procedure{DerivePoint}{$ Slope, Point, Length, IsSideLine $}

        \State init. $ a $, $ b $
        \Switch{$Slope$}
            \Case{$0$}
                \If{$ IsSideLine = True $}
                    \State $ a.x \gets Point.x $
                \Else
                    \State $ a.x \gets Point.x + Length $
                \EndIf
                \State $ a.y \gets Point.y $
                \State $ b.x \gets Point.x - Length $,~$ b.y \gets Point.y $
            \EndCase
            \Case{$ \infty $ or $ -\infty $}
                \State $ a.x \gets Point.x $

                \If{$ IsSideLine = True $}
                    \State $ a.y \gets Point.y $
                \Else
                    \State $ a.y \gets Point.y + Length $
                \EndIf

                \State $ b.x \gets Point.x $,~$ b.y \gets Point.y - Length $
            \EndCase
            \Default
                \State $ dx = \sqrt{\frac{Length}{{1+Slope^2}}}$,~ $ dy = Slope \cdot dx $
                \If{$ IsSideLine = True $}
                    \State $ a.x \gets Point.x $,~$ a.y \gets Point.y $
                \Else
                    \State $ a.x \gets Point.x + dx $,~$ a.y \gets Point.y + dy $
                \EndIf

                \State $ b.x \gets Point.x - dx $,~$ b.y \gets Point.y - dy $
            \EndDefault

>>>>

\section{Cartograms with Features}

{
\begin{figure}[b!]
    \centering
    \includegraphics[width=\columnwidth]{example-image-a}
    \caption{An overview of \software. Figure TBA.}
    \label{fig:overview}
\end{figure}
}

\algoref{alg:UpdateNodePosition} describes the overview of \software. We first load and render the CCG geospatial boundaries, for each CCG we compute the centroid where we render a node with $ \mathcal{S} $ as its size. We then apply the Fast Node Overlap Removal (VPSC) algorithm \cite{dwyer2006fast} recursively to remove overlaps. We chose VPSC over other node overlap removal algorithms since VPSC is able to provide spread minimization and node movement minimization while maintaining a good level of global shape preservation \cite{chen2020Node}. During each iteration, we test node trajectories (See \algoref{alg:check river crossing}) and move nodes that cross a river back to their previous position. A procedure ends when 1) no overlaps are present; 2) no nodes cross a river. We then increase $ \mathcal{S} $ by one pixel and repeat the algorithm until the entire map canvas is filled.

@@

\begin{algorithm}[tb!]
    \caption{The procedure to update node positions by removing overlap and prevent nodes from crossing rivers.}\label{alg:UpdateNodePosition}
    \textbf{Global variables:} \\
    $ \mathbcal{L} \gets $ a list of nodes representing CCGs \\
    $ \mathcal{S} \gets $ the current size of all nodes \\
    $ \mathcal{S}_{max} \gets $ the maximum size of a node \\
    $ \mathcal{I} \gets $ the maximum number of iterations before a stalemate \\

    \textbf{Local variables:} \\
    $ \mathbcal{L}_{vpsc} \gets $ a list of nodes representing CCGs after running VPSC \\
    $ cross \gets $ the boolean for whether a node crossed a river \\

    \begin{algorithmic}[1]
        \Procedure{UpdateNodePosition}{}
        \While{$ \mathcal{S} < \mathcal{S}_{max} $}

        \State $ \mathbcal{L}_{vpsc} \gets $ RunVPSC ($ \mathbcal{L} $)

            \ForEach {$ \mathbcal{L}^i_{vpsc} \in \mathbcal{L}_{vpsc} $}

                \If {TranslateNode ($ \mathbcal{L}^i, \mathbcal{L}^i_{vpsc} $)}
                    \State $ cross \gets $ \Call{TestRiverCross}{$ \mathbcal{L}^i, \mathbcal{L}^i_{vpsc} $}

                    \If{$ cross = True $}
                        \State $ \mathcal{I} \gets \mathcal{I} + 1 $

                        \If{$ \mathcal{I} < \mathcal{I} $}
                            \State \Call{TranslateNode}{$ \mathbcal{L}^i $}
                        \Else
                            \State \Call{DeriveCorridor}{$ \mathbcal{L}^i, \mathbcal{L}^i_{vpsc} $}
                        \EndIf

@@

        \State $ \mathcal{S} \gets \mathcal{S} ++$

@@

\begin{algorithm}[tb!]
    \caption{The procedure to update a single node's position.}\label{alg:move position}

    \textbf{Input:} \\
    $ \mathcal{N} \gets $ required, the node \\
    $ \mathcal{N}_{vpsc} \gets $ optional, the node after running VPSC \\

    \textbf{Output:} \\
    A boolean for whether the node's position is updated. \\

    \textbf{Local variables:} \\
    $ x,y \gets $ variables used to store coordinates \\

    \begin{algorithmic}[1]
        \Procedure{TranslateNode}{$ \mathcal{N}, \mathcal{N}_{vpsc} $}
        \If{$ \mathcal{N}.x = \mathcal{N}_{vpsc}.x $ and $ \mathcal{N}.y = \mathcal{N}_{vpsc}.y $}
            \item[] \Comment{Node position remains unchanged}
            \State \Return{$ False $}
        \EndIf

        \If{$ \mathcal{N}_{vpsc}$ is undefined } 
        \item[] \Comment{When optional parameter $ \mathcal{N}_{vpsc} $ is undefined}
        \item[] \Comment{Move the node back to its previous position}
            \State $ x \gets \mathcal{N}.previous.x,~y \gets \mathcal{N}.previous.y $
        \Else
            \State $ x \gets \mathcal{N}_{vpsc}.x,~y \gets \mathcal{N}_{vpsc}.y $
            \State $ \mathcal{N}.history.push(x, y) $
        \EndIf

        \State $ \mathcal{N}.x \gets x $
        \State $ \mathcal{N}.y \gets y $
        \State \Return{$ True $}
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

@@

\subsection{River Cross Testing}

We use rivers as topological boundaries and prevent nodes from crossing them. When a node's position changes, we test if the node's trajectory intersects any segment of a river. See \algoref{alg:check river crossing}. A bounding box intersection test can be performed to reduce the number of line intersection detections required.

% \begin{noindent}
\begin{algorithm}[tb!]
    \caption{The procedure to test if a node crosses a river.}\label{alg:check river crossing}
    \textbf{Input:} \\
    $ \mathcal{N} \gets $ required, the node \\
    $ \mathcal{N}_{vpsc} \gets $ required, the node after running VPSC \\

    \textbf{Output:} \\
    A boolean for whether the node crosses a river. \\

    \textbf{Global variables:} \\
    $ \mathbcal{R} \gets $ a list of river edges \\

    \textbf{Local variables:} \\
    $ b_\mathcal{N}, b_{\mathcal{R}^i} \gets $ the bounding boxes for $\mathcal{N}$ and $\mathcal{R}^i$ \\
    $ b_{intersect} \gets $ a boolean for whether $b_\mathcal{N}$ intersects $b_\mathcal{R}^i$ \\
    $ e_\mathcal{N}, b_{\mathcal{R}^i} \gets $ the edges for $\mathcal{N}$ and $\mathcal{R}^i$ \\
    $ e_{intersect} \gets $ a boolean for whether $e_\mathcal{N}$ intersects $e_\mathcal{R}^i$ \\

    \begin{algorithmic}[1]
        \Procedure{TestRiverCross}{$ \mathcal{N}, \mathcal{N}_{vpsc} $}

        \ForEach{$ \mathcal{R}^i \in \mathbcal{R} $}
            \State $ b_{\mathcal{N}} \gets $ GenerateBoundingBox ($ \mathcal{N}, \mathcal{N}_{vpsc} $)
            \State $ b_{\mathcal{R}^i} \gets $ GenerateBoundingBox ($ \mathcal{R}^i $)
            \State $ b_{intersect} \gets $ TestBoundingIntersection ($ b_{\mathcal{N}}, b_{\mathcal{R}^i} $)

            \If{$ b_{intersect} = True $}
                \State $ e_{\mathcal{N}} \gets $ GenerateEdge ($ \mathcal{N}, \mathcal{N}_{vpsc} $)
                \State $ e_{\mathcal{R}^i} \gets $ GenerateEdge ($ \mathcal{R}^i $)
                \State $ e_{intersect} \gets $ TestEdgeIntersection ($ e_\mathcal{N}, e_{\mathcal{R}^i} $)

                \If{$ e_{intersect} = True $}
                    \State \Return{$ True $}
                \EndIf

                \EndIf
        \EndFor

        \State \Return{$ False $}
        \EndProcedure
    \end{algorithmic}
\end{algorithm}
%\end{noindent}

@@

As the VPSC always tries to produce an optimal node layout where node spread and movement are minimized, a node might be repeatedly translated between two positions due to the lack of available space, creating a stalemate situation, as shown in \figref{fig:stalemate}. In order to create space for VPSC to generate a new layout, we introduce a user-adjustable parameter, $ \mathcal{I} $, to test for stalemate. If a node is translated between two positions for more than $ \mathcal{I} $ iterations, it is considered to be in a stalemate. A stalemate triggers a corridor derivation which moves all nodes within the corridor by $ \mathcal{D} $ set by the user (See \figref{fig:corridor} and \algoref{alg:derive corridor}).

{
\begin{figure}[tb!]
    \centering
    \includegraphics[width=\columnwidth]{figure/stalemate.png}
    \caption{A stalemate situation is when a node repeatedly translates between two positions (A and B) for more than $\mathcal{I}$ iterations. }
    \label{fig:stalemate}
\end{figure}
}

{
\begin{figure}[tb!]
    \centering
    \includegraphics[width=\columnwidth]{figure/corridor.png}
    \caption{When a stalemate occurs, a corridor (black rectangular) is derived based on the current (A) and previous (B) positions of the node that crossed a river, as described in \algoref{alg:derive corridor}. All nodes within the corridor are moved by $ \mathcal{D} $ in the direction of $ \protect\overrightarrow{BA\strut} $. }
    \label{fig:corridor}
\end{figure}
}

@@

\begin{algorithm}[tb!]
    \caption{The procedure to derive a corridor to move included nodes. We use a SVG canvas, where the point of origin (0,0) is located at the top left corner, with the x-axis extending to the right and the y-axis extending to the bottom, there is no negative axes.}\label{alg:derive corridor}

    \textbf{Input:} \\
    $ \mathcal{N} \gets $ required, the node current state \\
    $ \mathcal{N}_{p} \gets $ required, the node's previous state \\

    \textbf{Output:} \\
    A boolean for whether the node's position is updated. \\

    \textbf{Global variables:} \\
    $ \mathcal{C}_l \gets $ the length of a corridor \\
    $ \mathcal{C}_w \gets $ the width of a corridor \\
    $ \mathcal{D} \gets $ the distance to move nodes within the corridor \\
    $ \mathcal{S} \gets $ the current size of all nodes \\

    \textbf{Local variables:} \\
    $ slope \gets $ the slope of $ \overrightarrow{\mathcal{N}\mathcal{N}_{p}\strut} $ \\
    $ e_{p} \gets $ an edge perpendicular at $ \mathcal{N} $ on $ \overrightarrow{\mathcal{N}\mathcal{N}_{p}\strut} $ \\
    $ e_{s^1} \gets $ a perpendicular edge at the start of $ e_{p} $ \\
    $ e_{s^2} \gets $ a perpendicular edge at the end of $ e_{p} $ \\
    $ corridor \gets $ a rectangular formed by $ e_p $, $ e_{s^1} $ and $ e_{s^2} $ \\
    $ n_{in} \gets $ a node located inside $ corridor $ \\
    $ e_{d} \gets $ an edge connecting $ \mathcal{N} $ and its destination \\

    \begin{algorithmic}[1]
        \Procedure{DeriveCorridor}{$ \mathcal{N}, \mathcal{N}_{p} $}
            \State TranslateNode ($ \mathcal{N} $)

            \State $ \Delta x \gets \mathcal{N}.x - \mathcal{N}_{p}.x,~ \Delta y \gets \mathcal{N}.y - \mathcal{N}_{p}.y $

@@

            \State $ e_{p} = $ \Call{DeriveEdge}{$ \frac{-1}{ slope } $, $ \mathcal{N} $, $ \mathcal{C}_w $}
            \State $ e_{s^1} = $ \Call{DeriveEdge}{$ slope $, $ e_{p}.a $, $ \mathcal{C}_l $, $ True $}
            \State $ e_{s^2} = $ \Call{DeriveEdge}{$ slope $, $ e_{p}.b $, $ \mathcal{C}_l $, $ True $}
            \State $ corridor \gets
                \begin{bmatrix}
                    e_{p}.a &
                    e_{p}.b \\

                    e_{s^1}.a &
                    e_{s^2}.b \\
                \end{bmatrix} $

            \ForEach{$ {n}_{in}^i $ inside $ corridor $}
                \State $ e_{d} = $ \Call{DeriveEdge}{$ slope $, $ {n}_{in}^i $, $ \mathcal{D} $}

                \State TranslateNode ($ {n}_{in}^i, e_{d}.b $)
            \EndFor
        \EndProcedure
    \end{algorithmic}
\end{algorithm}

\begin{algorithm}[tb!]
    \caption{The procedure to derive an edge based on a slope and a point.}\label{alg:derive corridor point}
    \textbf{Input:} \\
    $ Slope \gets $ the slope used to derive an edge \\
    $ Point \gets $ the point used to derive an edge \\
    $ Length \gets $ the length of the derived edge \\
    $ IsSideEdge \gets $ the boolean used to determine the edge's type \\

    \textbf{Output:} \\
    An edge that passes through a given point with the given slope. \\

    \textbf{Local variables:} \\
    $ a,b \gets $ variables used to store coordinates \\
    $ \Delta x, \Delta y \gets $ the differences in $ x, y $ \\

    \begin{algorithmic}[1]
        \Procedure{DeriveEdge}{$ Slope, Point, Length, IsSideEdge $}

        \Switch{$Slope$}
            \Case{$0$}
                \If{$ IsSideEdge = True $}
                    \State $ a.x \gets Point.x $
                \Else
                    \State $ a.x \gets Point.x + Length $
                \EndIf
                \State $ a.y \gets Point.y $
                \State $ b.x \gets Point.x - Length $,~$ b.y \gets Point.y $
            \EndCase
            \Case{$ \infty $ or $ -\infty $}
                \State $ a.x \gets Point.x $

                \If{$ IsSideEdge = True $}
                    \State $ a.y \gets Point.y $
                \Else
                    \State $ a.y \gets Point.y + Length $
                \EndIf

                \State $ b.x \gets Point.x $,~$ b.y \gets Point.y - Length $
            \EndCase
            \Default
                \State $ \Delta x = \sqrt{\frac{Length}{{1+Slope^2}}}$,~ $ \Delta y = Slope \cdot \Delta x $
                \If{$ IsSideEdge = True $}
                    \State $ a.x \gets Point.x $,~$ a.y \gets Point.y $
                \Else
                    \State $ a.x \gets Point.x + \Delta x $,~$ a.y \gets Point.y + \Delta y $
                \EndIf

                \State $ b.x \gets Point.x - \Delta x $,~$ b.y \gets Point.y - \Delta y $
            \EndDefault

### 2022-02-18 paper/section/related.tex
draft: update

This section introduces the definition of cartogram types, describes relevant applications of cartograms, and provides a brief overview of some real-world applications of cartogram-based visualizations.

@@

Dorling cartograms are non-contiguous and do not preserve geography and topology. A Dorling cartogram is statistically accurate, regions are represented by circles and the statistic of interest is represented by the circle size \cite{dorling2011Area}. 
In Demers Cartograms, a variant of Dorling, squares are used instead to capture a certain level of topology \cite{cano2015Mosaic}. Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlaps.

Rectangular cartograms are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular}. Depending on the variant, a rectangular cartogram may trade between statistical and topological accuracy.

Mosaic cartograms are contiguous and sacrifice statistical accuracy to preserve some level of geographical accuracy \cite{cano2015Mosaic}. Some variants are able to preserve topological accuracy as well.

\citea{warf2008Geography} use a Dorling cartogram to represent the religious diversity in the United States. \citea{sun2010Effectiveness} visualize 1996 US election data and 2005 China population date using Dorling, Mosaic, and contiguous cartograms. \citea{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal. \citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China. \citea{tong2018cartograms} use a Demers cartogram to visualize health-related date by regions in England, the work introduces a novel technique to remove overlaps of squares based on topological features, aiming to improve both geographical and topological accuracy. \citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US. See Table \ref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.

{
\renewcommand{\arraystretch}{1.5}
\begin{table*}[!tb]
	\centering
	\resizebox{\textwidth}{!}{
		\begin{tabulary}{\textwidth}{|*{4}{l|}r|c|}
			\hhline{~|*{4}{-}|~}
			\multicolumn{1}{c|}{\textbf{Literature}} &
			\textbf{Title} &
			\textbf{Cartogram Type(s)} &
			\textbf{Geographic Region(s)} &
			\textbf{Number of Nodes} &
			\multicolumn{1}{c}{\textbf{Year}} \\
			\hline

@@

		\end{tabulary}
	}
	\caption{}
	\label{table:region vs node}
\end{table*}
}

@@

Cartogram has become a popular choice of visualization in covering various topics by the media, in order to present an unbiased and complete picture of the news. \citea{newman2016Election} generates contiguous cartograms to present US Election results in 2012 and 2016. The Washington Post adopts cartograms to visualize the US overseas economic assistance, arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}. The National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}. \citea{sandberg2018Cartogram} reports the 2018 US midterm Election with a Mosaic cartogram, the same approach is used to cover the 2020 US Election by the New York Times \cite{thelearningnetwork2020What} and Bloomberg \cite{mccartney20202020}.

>>>>

This section introduces the characteristics of various cartogram styles, describes relevant applications of cartograms, and provides a brief overview of some real-world applications of cartogram-based visualizations.

@@

{
\renewcommand{\arraystretch}{1.5}
\begin{table*}[!htb]
	\centering
	\resizebox{\textwidth}{!}{
		\begin{tabulary}{\textwidth}{|*{4}{l|}r|c|}
			\hhline{~|*{4}{-}|~}
			\multicolumn{1}{c|}{\textbf{Literature}} &
			\textbf{Title} &
			\textbf{Cartogram Type(s)} &
			\textbf{Geographic Region(s)} &
			\textbf{Number of Nodes} &
			\multicolumn{1}{c}{\textbf{Year}} \\
			\hline

@@

		\end{tabulary}
	}
	\caption{Related work with cartogram-based visualizations. \textbf{Cartogram type} is the type of cartogram used. \textbf{Geographic region} is the geographic region depicted by the cartogram. \textbf{Number of nodes} is the number of nodes depicted in the cartogram. }
	\label{table:region vs node}
\end{table*}
}

\thumbnail{dorling}{
	Dorling cartograms are non-contiguous and do not preserve geography and topology. A Dorling cartogram is statistically accurate, regions are represented by circles and the data dimension of interest is represented by the circle size \cite{dorling2011Area}. In a Demers Cartogram, a variant of Dorling, squares are used instead to capture a certain level of topology \cite{cano2015Mosaic}. Dorling cartograms are unable to maintain topological accuracy as circles are often repositioned to remove overlap.
}

\thumbnail{rectangular}{
	Rectangular cartograms are contiguous and do not preserve geographical accuracy \cite{raisz1934Rectangular, vankreveld2004Rectangular}. Depending on the variant, a rectangular cartogram may trade off between statistical and topological accuracy.
}

\thumbnail{mosaic}{
	Mosaic cartograms usually use square or hexagonal tiles to depict regions, and are contiguous and sacrifice statistical accuracy to preserve some level of geographical accuracy \cite{cano2015Mosaic}. Some variants are able to preserve topological accuracy as well.
}

\citea{warf2008Geography} use a Dorling cartogram to represent the religious diversity in the United States. \citea{sun2010Effectiveness} depict 1996 US election data and 2005 Chinese population data using Dorling, Mosaic, and contiguous cartograms. \citea{gao2020Visualising} present a Dorling cartogram to illustrate COVID-19 infections in China. \citea{tong2018cartograms} use a Demers cartogram to visualize health-related date by regions in England, the work introduces a novel technique to remove overlap of squares based on topological features, aiming to improve both geographical and topological accuracy. \citea{nusrat2020Recognition} investigate the memorability of contiguous and Dorling cartograms by using multiple datasets including demographics, agriculture, and retail data in the US. See Table \ref{table:region vs node} for a list of literature that adopts cartograms for visualization with corresponding geographical regions and node counts.

@@

Cartograms are a popular choice of representation in covering various topics by the media, in order to present an unbiased and complete picture of the news. The Washington Post uses cartograms to visualize the US overseas economic assistance, arm sales (Mosaic) \cite{bearak2016Everything}, the 2016 US Election (contiguous) \cite{gamio2016Election}, and the Brexit Referendum (Mosaic) \cite{taylor2016What}. The National Geographic uses contiguous and Mosaic cartograms to analyze the 2016 US Election results \cite{miller2016Election}, the same topic is also covered by the Financial Times with a Dorling cartogram \cite{stabe2016Search}. \citea{cruz2017Adapted} adapts a Dorling cartogram with both contiguous and non-contiguous cartograms to represent the gender pay gap in Portugal. \citea{sandberg2018Cartogram} reports the 2018 US midterm Election with a Mosaic cartogram, the same approach is used to cover the 2020 US Election by the New York Times \cite{thelearningnetwork2020What} and Bloomberg \cite{mccartney20202020}.

% \subsection{Contiguous Applications}

% \citea{newman2016Election} generates contiguous cartograms to present US Election results in 2012 and 2016.

Compared to previous work on Demers Cartogram, we introduce a topological feature, a river, aiming to improve the interpretation, readability, and topological accuracy.
